---
title: "dataset_intro"
author: "Sarah Deussing"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading Packages
```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(caret)
library(ggplot2)
library(readr)
library(xgboost)
library(stringr)
library(lubridate)
```

## Read Dataset
```{r Read in data}
fraud <- read.csv("fraud_data.csv")
summary(fraud)
```

Data conversions needed:
- trans_data_trans time: separate into date + time

##Data Conversion
```{r}
fraud[c('trans_date', 'trans_time')] <- str_split_fixed(fraud$trans_date_trans_time, " ", 2)
fraud

fraud$trans_date <- mdy(fraud$trans_date)
fraud
```
The response variable is 'is_fraud'. 1 denotes fraud, 0 denotes not fraud.

The five explanatory variables we are looking at are:
  - Category
  - Amount
  - Merchant
  - Time (Time of day or time of year)
  - Age from DOB
dfghjkl

Run some exploratory analysis
- maybe PCA??
- make a plot of fraud by longtidue/latitude coordinates (sender and merchant)
- some other plots

Run logistic regression
- split into training + test sets
- run lasso logistic regression model
- look at variance, accuracy, etc.
- if dataset is unbalanced, balance and run again

Time Series Modeling
- amount over time
